---
output:
  html_document:
    self_contained: no
    dev: png
---

```{r, include = FALSE, echo=FALSE, fig.height= 7, fig.width=7}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height= 7, 
  fig.width=7
)
```

## Installation

The code can be installed in R from GitHub using the `remotes` (or `devtools`) package.  Please be aware that as development code, it may be broken at times. 
```{r remotes, eval=FALSE}
remotes::install_github("https://github.com/mbonert/cmpproviders")
```

## The provider factor
The provider is a known predictor of outcome in healthcare. In the quality improvement context: it is useful to understand the magnitude of differences and, more generally, the variability. 

## Funnel plots
Funnel plots are useful to understand variability and the magnitude of differences.  Properly constructed, they can be understood by most practicing physicians/surgeons with modest effort. 

This code borrows some ideas from the package "FunnelPlotR". The main differences to that package are: (1) this code uses the raw rate data (instead of the standardized ratio), (2) the confidence intervals the funnels represent can be adjusted, and (3) an arbitrary number of funnels can be plotted in the same figure.

### Example 1 (Length of Stay)
```{r, warning=FALSE, error=FALSE, message=FALSE, fig.height= 7, fig.width=7}
library(cmpproviders)
library(COUNT)
library(dplyr)
data(medpar)

print("Get number of patients by provider ...")
provider_count=as.data.frame(table(medpar$provnum))
colnames(provider_count) = c('Provider', 'Count')
provider_count

print("Purging providers with less than 25 patients ...")
requiredcases=25
medpar_trunc <- medpar[medpar$provnum %in% provider_count[provider_count$Count>=requiredcases,]$Provider,]

print("Number of patients by provider in purged data set ...")
provider_count_trunc=as.data.frame(table(medpar_trunc$provnum))
colnames(provider_count_trunc) = c('Provider', 'Count')
provider_count_trunc

print("Use '14' as the target length of stay ...")
target_los=14
medpar_trunc_gt_target_los <- medpar_trunc[medpar_trunc$los>target_los,]

medpar_trunc$gt_target_los=0
medpar_trunc[medpar_trunc$los>target_los,]$gt_target_los <- 1
prov_vs_missrate_table=table( medpar_trunc$provnum, medpar_trunc$gt_target_los)
num_prov=dim(prov_vs_missrate_table)[1]


prov_matrix = matrix(NA, num_prov, 3)
prov_matrix[,1]=prov_vs_missrate_table[,2]
prov_matrix[,2]=prov_vs_missrate_table[,1]+prov_vs_missrate_table[,2]
prov_matrix[,3]=prov_matrix[,1]/prov_matrix[,2]
colnames(prov_matrix) = c('Exceed LOS target', 'Patient count', 'Miss rate')
prov_matrix

x_var=prov_matrix[,2]
y_var=prov_matrix[,3]

provider_labels = provider_count_trunc[,1]

print("Creating funnel plot ...")
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9,99.9999,99.9999999), labels=provider_labels, addlabels = 1, x_label = "Number of Patients for Each Provider", y_label = "Number of Patients Exceeding Target LOS/Number of Patients", plot_title = "Length of Stay (LOS) Target Miss Rate by Provider and Volume", y_percent=FALSE)
```

### Example 2 (Death Rate)
```{r, warning=FALSE, error=FALSE, message=FALSE, fig.height= 7, fig.width=7}
library(cmpproviders)
library(COUNT)
library(dplyr)
data(medpar)

print("Number of patients by provider ...")
provider_count=as.data.frame(table(medpar$provnum))
colnames(provider_count) = c('Provider', 'Count')
provider_count

print("Purging providers with less than 20 patients ...")
requiredcases=20
medpar_trunc <- medpar[medpar$provnum %in% provider_count[provider_count$Count>=requiredcases,]$Provider,]

print("Number of patients by provider in purged data set ...")
provider_count_trunc=as.data.frame(table(medpar_trunc$provnum))
colnames(provider_count_trunc) = c('Provider', 'Count')
provider_count_trunc

medpar_trunc_died <- medpar_trunc[medpar_trunc$died==1,]

print("Number of died by provider in purged data set ...")
provider_count_trunc_died=as.data.frame(table(medpar_trunc_died$provnum))
colnames(provider_count_trunc_died) = c('Provider', 'Count')
provider_count_trunc_died

print("Death rate by provider ...")
y_var = provider_count_trunc_died[,2]/ provider_count_trunc[,2]
y_var

x_var = provider_count_trunc[,2]
provider_labels = provider_count_trunc[,1]

print("Creating funnel plots ...")

fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9), labels=provider_labels, addlabels = 1, x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume") 

print("Labelling one provider (030006) ...")
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9), labels=provider_labels, addlabels = 'LIST', labels_to_display = c("030006"), x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume") 

print("Labels only the outliers ...")
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9,99.9999), labels=provider_labels, addlabels = "OUTLIERS", x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Provider Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume with Outliers Labelled") 

```
