---
title: "Compare Providers using Funnel Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compare (Healthcare) Providers using Rate Data and Funnel Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo=FALSE, fig.height= 7, fig.width=7}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height= 7, 
  fig.width=7
)
```

### The provider factor
The provider is a known predictor of outcome in healthcare. In the quality improvement context: it is useful to understand the magnitude of differences and, more generally, the variability. 

### Funnel plots
Funnel plots are useful to understand variability and the magnitude of differences.  Properly constructed, they can be understood by most practicing physicians/surgeons with modest effort. 

### Example 1 (Length of Stay)
```{r, warning=FALSE, error=FALSE, message=FALSE, fig.height= 7, fig.width=7}
library(cmpproviders)
library(COUNT)
library(dplyr)
data(medpar)

# Get number of patients by provider ...
provider_count=as.data.frame(table(medpar$provnum))
colnames(provider_count) = c('Provider', 'Count')
provider_count

# Purging providers with less than 25 patients ...
requiredcases=25
medpar_trunc <- medpar[medpar$provnum %in% provider_count[provider_count$Count>=requiredcases,]$Provider,]

# Number of patients by provider in purged data set ...
provider_count_trunc=as.data.frame(table(medpar_trunc$provnum))
colnames(provider_count_trunc) = c('Provider', 'Count')
provider_count_trunc

# Use '14' as the target length of stay ...
target_los=14
medpar_trunc_gt_target_los <- medpar_trunc[medpar_trunc$los>target_los,]

medpar_trunc$gt_target_los=0
medpar_trunc[medpar_trunc$los>target_los,]$gt_target_los <- 1
prov_vs_missrate_table=table( medpar_trunc$provnum, medpar_trunc$gt_target_los)
num_prov=dim(prov_vs_missrate_table)[1]


prov_matrix = matrix(NA, num_prov, 3)
prov_matrix[,1]=prov_vs_missrate_table[,2]
prov_matrix[,2]=prov_vs_missrate_table[,1]+prov_vs_missrate_table[,2]
prov_matrix[,3]=prov_matrix[,1]/prov_matrix[,2]
colnames(prov_matrix) = c('Exceed LOS target', 'Patient count', 'Miss rate')
prov_matrix

x_var=prov_matrix[,2]
y_var=prov_matrix[,3]

provider_labels = provider_count_trunc[,1]

# Creating funnel plot ...
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9,99.9999,99.9999999), labels=provider_labels, addlabels = 1, x_label = "Number of Patients for Each Provider", y_label = "Number of Patients Exceeding Target LOS/Number of Patients", plot_title = "Length of Stay (LOS) Target Miss Rate by Provider and Volume", y_percent=FALSE)


```

### Example 2 (Death Rate)
```{r, warning=FALSE, error=FALSE, message=FALSE, fig.height= 7, fig.width=7}
library(cmpproviders)
library(COUNT)
library(dplyr)
data(medpar)

# Number of patients by provider ...
provider_count=as.data.frame(table(medpar$provnum))
colnames(provider_count) = c('Provider', 'Count')
provider_count

# Purging providers with less than 20 patients ...
requiredcases=20
medpar_trunc <- medpar[medpar$provnum %in% provider_count[provider_count$Count>=requiredcases,]$Provider,]

# Number of patients by provider in purged data set ...
provider_count_trunc=as.data.frame(table(medpar_trunc$provnum))
colnames(provider_count_trunc) = c('Provider', 'Count')
provider_count_trunc

medpar_trunc_died <- medpar_trunc[medpar_trunc$died==1,]

# Number of died by provider in purged data set ...
provider_count_trunc_died=as.data.frame(table(medpar_trunc_died$provnum))
colnames(provider_count_trunc_died) = c('Provider', 'Count')
provider_count_trunc_died

# Death rate by provider ...
y_var = provider_count_trunc_died[,2]/ provider_count_trunc[,2]
y_var

x_var = provider_count_trunc[,2]
provider_labels = provider_count_trunc[,1]

# Creating funnel plots ...

fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9), labels=provider_labels, addlabels = 1, x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume") 

# Labelling one provider (030006) ...
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9), labels=provider_labels, addlabels = 'LIST', labels_to_display = c("030006"), x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume") 

# Labels only the outliers ...
fp=funnel2cmpproviders(x_var, y_var, limits=c(95,99.9,99.9999), labels=provider_labels, addlabels = "OUTLIERS", x_label = "Provider Volume (Patients Cared for by Provider)", y_label = "Provider Death Rate (Deaths by Provider/Patients Cared for by Provider)", plot_title = "Death Rate by Provider and Volume with Outliers Labelled") 

```
